Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-08-14 doc/manuscript/manuscript.Rmd
white mold nurs -> white mold screening nurs

In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S. sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia. 
Because the white mold nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S. sclerotiorum* populations from white mold nurseries were representative of the fields within the same region. 
As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
By investigating these aims, we hope to effectively describe the population dynamics of *S. sclerotiorum* and make available our database of isolates for use in future dry bean breeding efforts.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The white mold nurseries in different regions are indicated by wmn (white mold nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).

@@

The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)). 
In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).

@@

Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between White Mold Nurseries and other sources. The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold nursery or not. Bold $\Phi$ values indicate significant values (P < 1.00e^-4^). Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI. S.S. = Sum of Squares, d.f. = degrees of freedom.

@@

<!-- 
 - Big Picture reminder of importance
   + This is an extension of a previous work [@otto-hanson2011variation]
   + White Mold Nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
   + It's unclear how genetic structure can influence
 - Critical Analysis of Major Findings
   + This is the largest genetic analysis of *S. sclerotiorum* to date
   + Similar to the findings of @otto-hanson2011variation, we found that
     Region has a strong influence on both population structure and
     aggressiveness.
   + North Dakota appeared
 - Additional Findings and how they fit into the literature
 - Limitations
 - Future Directions
 - Overall Conclusion and Impact
-->

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level. 
We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the white mold nurseries can be effective at screening for potential resistant lines within growing regions.

@@

The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
When we assessed AMOVA using only populations with equally representative white mold screening nursery isolates and producer field isolates, we see the same pattern, except the $\Phi$ values decrease by nearly half (Table \@ref(tab:amova)). 
When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the patterns of differentiation are largely driven by isolates from Mexico and California. 
Isolates from these populations had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
All other populations in comparison (except France) has reassignment probabilities of < 0.5, which is reflected in the failure of the first two discriminant functions to separate these populations (Fig. \@ref(DAPC)b).
DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple]. 
We found no evidence of private alleles in the Mexican and Californian isolates, suggesting that this pattern separating these populations from the others was not an artifact.

@@

Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the white mold nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
This may eventually result in facilitating outcrossing and recombination.

@@

There were isolates that shared the same MLG belonging to different MCGs.
For instance MLG 25 consisted of 27 isolates in 5 MCGs.
This phenomenon is suggestive of a recombination by sexual outcrossing [@njambere2014genotypic; @sexton2006population].  In general high level of MLG diversity; and lack of association of MCG and microsatellite genotype are evidence of *S. sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices [@njambere2014genotypic].
Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
Linkage disequilibrium analysis of populations in white mold nurseries indicated the possibility of sexual outcrossing in some regions (ex.
CA, FR, OR and WA) during certain sampling years while most locations indicated clonal reproduction.
For example wmn of WA was sampled in 2003, 2004, 2005, and 2008 but only 2008 isolates had evidence of outcrossing.
Similar to our results many studies have reported primarily clonal reproduction of *S. sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA [@aldrich-wolfe2015genetic]; cabbage population in North Carolina [@cubeta1997clonality]; Canadian canola and North Carolina cabbage populations [@kohli1998random]; and 12 populations from five crops in England and Wales [@clarkson2013population].
Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal [@cubeta1997clonality].
@atallah2004high studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
When all isolates were combined, there was significant linkage disequilibrium and hence the population was highly clonal.
Our study supports this phenomenon since linkage disequilibrium analysis revealed evidence of sexual outcrossing only in CA when populations were based on region.
However, when populations were sub divided according each year/region, we could detect evidence of sexual out crossing in some populations.
One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature [@atallah2004high].

>>>>

In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S. sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia. 
Because the white mold screening nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold screening nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
By investigating these aims, we hope to effectively describe the population dynamics of *S. sclerotiorum* and make available our database of isolates for use in future dry bean breeding efforts.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or white mold screening nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The white mold screening nurseries in different regions are indicated by wmn (white mold screening nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).

@@

The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold screening nurseries and non-white mold screening nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)). 
In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).

@@

Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries and other sources. The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold screening nursery or not. Bold $\Phi$ values indicate significant values (P < 1.00e^-4^). Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI. S.S. = Sum of Squares, d.f. = degrees of freedom.

@@

<!-- 
 - Big Picture reminder of importance
   + This is an extension of a previous work [@otto-hanson2011variation]
   + white mold screening nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
   + It's unclear how genetic structure can influence
 - Critical Analysis of Major Findings
   + This is the largest genetic analysis of *S. sclerotiorum* to date
   + Similar to the findings of @otto-hanson2011variation, we found that
     Region has a strong influence on both population structure and
     aggressiveness.
   + North Dakota appeared
 - Additional Findings and how they fit into the literature
 - Limitations
 - Future Directions
 - Overall Conclusion and Impact
-->

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level. 
We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the white mold screening nurseries can be effective at screening for potential resistant lines within growing regions.

@@

The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold screening nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
When we assessed AMOVA using only populations with equally representative white mold screening nursery isolates and producer field isolates, we see the same pattern, except the $\Phi$ values decrease by nearly half (Table \@ref(tab:amova)). 
When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the patterns of differentiation are largely driven by isolates from Mexico and California. 
Isolates from these populations had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
All other populations in comparison (except France) has reassignment probabilities of < 0.5, which is reflected in the failure of the first two discriminant functions to separate these populations (Fig. \@ref(DAPC)b).
DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple]. 
We found no evidence of private alleles in the Mexican and Californian isolates, suggesting that this pattern separating these populations from the others was not an artifact.

@@

Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the white mold screening nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
This may eventually result in facilitating outcrossing and recombination.

@@

There were isolates that shared the same MLG belonging to different MCGs.
For instance MLG 25 consisted of 27 isolates in 5 MCGs.
This phenomenon is suggestive of a recombination by sexual outcrossing [@njambere2014genotypic; @sexton2006population].  In general high level of MLG diversity; and lack of association of MCG and microsatellite genotype are evidence of *S. sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices [@njambere2014genotypic].
Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
Linkage disequilibrium analysis of populations in white mold screening nurseries indicated the possibility of sexual outcrossing in some regions (ex.
CA, FR, OR and WA) during certain sampling years while most locations indicated clonal reproduction.
For example wmn of WA was sampled in 2003, 2004, 2005, and 2008 but only 2008 isolates had evidence of outcrossing.
Similar to our results many studies have reported primarily clonal reproduction of *S. sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA [@aldrich-wolfe2015genetic]; cabbage population in North Carolina [@cubeta1997clonality]; Canadian canola and North Carolina cabbage populations [@kohli1998random]; and 12 populations from five crops in England and Wales [@clarkson2013population].
Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal [@cubeta1997clonality].
@atallah2004high studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
When all isolates were combined, there was significant linkage disequilibrium and hence the population was highly clonal.
Our study supports this phenomenon since linkage disequilibrium analysis revealed evidence of sexual outcrossing only in CA when populations were based on region.
However, when populations were sub divided according each year/region, we could detect evidence of sexual out crossing in some populations.
One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature [@atallah2004high].
